hbm_paper_results

### 2024-01-04 paper_results.Rmd
cleaning up stat tests for main analyses

Test for differences in abs. error distributions across combat configurations using nonparametric kruskal-wallace test (instead of anova. good that it's nonparametric b/c non-normal distributions BUT also the data is paired!). **confirm with andrew**
Within each phenotype, iterate paired t test of ranks between each combo of combat dfs (nonparametric per discussion with Andrew). (see Zimmerman & Zumbo, 1993). Using custom functions from `ranked_welch_tests.R`.

**need to confirm w andrew then update output table formatting as needed - skipping for now**

@@

```{r}
sex.bias.t.tests(subj_sum.df)
```

@@

```{r}
pred.csvs <- list.files(path = data_path, pattern = "_variance.csv", full.names = TRUE)
combined_var <- data.frame()

@@

```{r}
#var plot
var.plot.df <- combined_var %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume") %>%
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset, pheno_cat)

@@

```{r}
basic_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_basic"
raw_files <- list.files(path = basic_data_path, pattern = "_centiles.csv", full.names = TRUE)

@@

```{r}
perm_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_perm_csvs"
raw_files <- list.files(path = perm_data_path, pattern = "_diffs.csv", full.names = TRUE)

@@

>>>>

Within each phenotype, iterate paired t test of ranks between each combo of combat dfs (nonparametric per discussion with Andrew). (see Zimmerman & Zumbo, 1993). Using custom functions from `ranked_welch_tests.R`.

@@

```{r}
sex.bias.t.tests(subj_sum.df)
sex.bias.t.tests(subj_sum.df, to_test="mean_cent_diff")
```

### Extremeness-Biases

Group at 0-0.25, >0.25 - .75 <, 0.75-1

```{r}
subj_sum.df <- subj_sum.df %>%
  mutate(cent_cat = factor(case_when(mean_centile <= 0.25 ~ "low",
                              mean_centile > 0.25 & mean_centile < 0.75 ~ "mid",
                              mean_centile >= 0.75 ~ "high"), levels= c("low", "mid", "high"), ordered=TRUE),
         z_cat = as.factor(case_when(mean_z <= -1 ~ "low",
                              mean_z > -1 & mean_z < 1 ~ "mid",
                              mean_z >= 1 ~ "high")))
```

```{r}
subj_sum.df %>%
  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
  geom_point(alpha=0.3) +
  geom_smooth() +
  facet_grid(dataset ~ sim.site, scales="free_y")
```

```{r}
subj_sum.df %>%
  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
  ggplot(aes(x=cent_cat, y=mean_cent_diff, color=sex)) +
  geom_violin() +
  facet_wrap(~dataset, scales="free_y")

subj_sum.df %>%
  #dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
  ggplot(aes(x=cent_cat, y=mean_cent_diff)) +
  geom_violin() +
  facet_wrap(~dataset, scales="free_y")
```

@@

```{r}
pred.csvs <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_variance.csv", full.names = TRUE)
combined_var <- data.frame()

@@

```{r}
#var plot
var.plot.df <- combined_var %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume") %>%
  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset, pheno_cat)

@@

```{r}
raw_files <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_centiles.csv", full.names = TRUE)

@@

```{r}
perm_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_perm"
raw_files <- list.files(path = perm_data_path, pattern = "_diffs.csv", full.names = TRUE)

@@

```{r}
for (l in unique(ratio_subj.df$dataset)){
plot <- ratio_subj.df %>%
  dplyr::filter(dataset == l) %>%
  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
  geom_point(alpha=0.3) +
  geom_smooth() +
  facet_grid(sim.site ~ prop) +
  ggtitle(l)
print(plot)
}
```
